\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-06-08 nampi.tex
shortened things

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

@@

  \begin{tabular}{l}
    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
    \popp{$f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))}}\\
    \hspace{\helpSize}\emph{($f_6$: rational function)}\\

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
\end{table*}
Because any learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on problems that people solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  %% Even with a good DSL, program search may be
%% intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
%% so we train a
%% neural network to guide the search procedure.

@@

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$. %% \footnote{For example, for string editing,
%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
%% and 0 otherwise.}
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL, written $\mathcal{D}$,
which is a generative model over programs, written $\probability[p |\mathcal{D}]$.
We frame our goal as maximum a posteriori (MAP) inference of $\mathcal{D}$ given $X$:
$$
\argmax_{\mathcal{D}}\probability[\mathcal{D}]\prod_{x\in X}\sum_p\probability[x|p]\probability[p|\mathcal{D}]
$$
Even evaluating the above objective is intractable due to the sum over all programs,
and so we instead marginalize over finite sets of programs, called \textbf{frontiers}, one for each task $x\in X$, written $\mathcal{F}_x$, giving the following lower bound on the marginal, which we call $\lowerBound$:
\begin{equation}\label{lowerBound}
  \lowerBound  \triangleq\probability[\mathcal{D}]\prod_{x\in X}\sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D}]
\end{equation}
We alternate maximization w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$ (\textbf{Wake}) and $\mathcal{D}$  (\textbf{Sleep-G}):

\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the programs.} Here $\mathcal{D}$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D}]$
is large.
Here we use a simple and generic enumerative program synthesis algorithm.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\mathcal{D}$.
To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.

@@

\subsection{Programs that manipulate sequences}\label{sequences}
We apply \system to list processing %% (Section~\ref{listSection})
and text editing, %% (Section~\ref{textSection})
using a bidirectional GRU~\cite{cho2014learning} for
the recognition model, and initially providing the system with generic sequence manipulation primitives:
\code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

%\subsubsection{List Processing}\label{listSection}
\textbf{List Processing:} Synthesizing programs that manipulate data structures is a widely studied
problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of learning functions that
manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples.
%% Our data set is interesting in three major ways: many of the tasks
%% require complex solutions; the tasks were not generated from some
%% latent DSL, and the agent must learn to solve these complicated
%% problems from only 236 tasks.
%% Our data set assumes arithmetic operations as well as sequence operations,
%% so we additionally provide our system with the following arithmetic
%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
%% \code{is-prime}.
%% evaluating  on a 50/50 test/train split.
In solving these tasks, the system
composed 38 new subroutines, and rediscovered the higher-order
function \code{filter} ($f_1$ in Table~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
\begin{figure}[b]\centering
  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
%    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
%    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-mod-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}
\end{figure}

@@

\subsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial or rational function.  The recognition model is a
convnet that observes an image of the target function's
graph (Fig.~\ref{functions}) --- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can  look at a graph and predict
what kind of function best explains it.
These programs can contain real numbers,
which we set with gradient descent,
and penalize continuous parameters by
incorporating a BIC penalty into the likelihood model $\probability[x|p]$.
%% A key difficulty, however, is
%% that these problems are best solved with programs containing real
%% numbers.  Our solution to this difficulty is to enumerate
%%  programs with real-valued parameters, and then fit those
%% parameters by automatically differentiating through the programs the
%% system writes and use gradient descent to fit the parameters.
%% We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
%% and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
  \begin{minipage}[c]{0.3\columnwidth}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}\\
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}%\includegraphics[width = \functionSize]{figures/functions/181.png}
  %\includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
    \includegraphics[width = \functionSize]{figures/functions/149.png}
  \end{minipage}
  \begin{minipage}[c]{0.69\columnwidth}    
    \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}
        \end{minipage}
\end{figure}
We learn a DSL containing 13 new functions,
mainly templates for polynomials of different orders or ratios of polynomials.
The model also learns to find programs minimizing the number of continuous parameters ---
learning to represent linear functions with 
\code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing.

@@

We ran the following experiment: \system was provided with a subset of
the 1959 Lisp primitives, and tasked with solving 22 functional
programming exercises.  After running for 93 hours on 64 CPUs, our
algorithm solves these exercises, along the way assembling a DSL
with a modern repertoire of
functional programming idioms and subroutines, including \code{map},
\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
and simple arithmetic operations like 
building lists of natural numbers between an interval.
%% We did not use the recognition model for this experiment:
%% a bottom-up pattern recognizer is of little use
%% for acquiring this abstract knowledge
%% from less than two dozen problems.

@@

\subsection{Quantitative Results}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models 
more closely matching the domain.

@@

\begin{figure*}\centering
  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
  \caption{Learning curves for \system both with (\orange{in orange}) and without
    (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved w/ 10m timeout. Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}
\end{figure*}

\begin{comment}
\section{Related Work}
 Our work is far from the first for learning to learn programs,
 an idea that goes back to Solomonoff~\cite{solomonoff1989system}:

 \noindent \textbf{Deep learning:} Much recent work in the ML community has
 focused on creating neural networks that regress from
 input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. %This family of work is closest to our RF/DC baseline.
\systemEnding's recognition model draws heavily from this line of work, particularly from~\cite{menon2013machine}.
We see these prior works as operating in a different regime: typically, they train with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
 Our work   considers a weakly-supervised regime where ground truth programs are not provided and
 the agent must learn from at most a few hundred tasks,
 which is facilitated by our ``Helmholtz machine'' style recognition model.%(Sec.~\ref{recognitionSection}).

 \noindent \textbf{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.

\noindent\textbf{Bayesian Program
 Learning:} Our work is an instance of
 Bayesian Program
 Learning (BPL; see~\citep{lake2015human,Dechter:2013:BLV:2540128.2540316,ellis2016sampling,DBLP:conf/icml/LiangJK10}). Previous BPL systems have largely assumed a fixed DSL (but see~\cite{DBLP:conf/icml/LiangJK10}),
 and our contribution here is a general way of doing BPL with less hand-engineering of the DSL.
  \end{comment} 
\section{Discussion}

%\paragraph{Outlook.}
We contribute an algorithm, \systemEnding, that learns to program by
bootstrapping a DSL with new domain-specific primitives that the algorithm
itself discovers, together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration of top-down
symbolic representations and bottom-up neural nets --- both of them learned
--- can help make program induction systems more generally useful for AI. Many
directions remain open.
%\paragraph{Future.}
Two immediate goals are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the generic methods used here.
Another direction is DSL meta-learning: can we find a
\emph{single} universal primitive set that could effectively bootstrap DSLs for
new domains, including the domains considered here,  but also many others?

>>>>

\usepackage{dblfloatfix}    % To enable figures at the bottom of page
% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

@@

  \begin{tabular}{l}
    \popp{$f_0($\code{x}$)\,=\,$\code{(+ x real)}}\\
    \popp{$f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} }\\
    \popp{$f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}}\\
    \popp{$f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}}\\
    \popp{$f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}}\\
    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
    \popp{$f_5($\code{x}$)\,=\,$\code{(/ real x)}}\\
    \popp{$f_6($\code{x}$)\,=\,$\code{($f_5$ ($f_0$ x))}}\\
    \hspace{\helpSize}\emph{($f_6$: rational function)}\\

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}\vspace{-0.5cm}
\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}\vspace{-0.5cm}
\end{table*}
Because any learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on problems that people solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  %% Even with a good DSL, program search may be
%% intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
%% so we train a
%% neural network to guide the search procedure.

@@

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$. %% \footnote{For example, for string editing,
%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
%% and 0 otherwise.}
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL, written $\mathcal{D}$,
which is a generative model over programs, written $\probability[p |\mathcal{D}]$.
We frame our goal as maximum a posteriori (MAP) inference of $\mathcal{D}$ given $X$:
$$
\argmax_{\mathcal{D}}\probability[\mathcal{D}]\prod_{x\in X}\sum_p\probability[x|p]\probability[p|\mathcal{D}]
$$
Even evaluating the above objective is intractable due to the sum over all programs,
and so we instead marginalize over finite sets of programs, called \textbf{frontiers}, one for each task $x\in X$, written $\mathcal{F}_x$, giving the following lower bound on the marginal, which we call $\lowerBound$:
\begin{equation}\label{lowerBound}
  \lowerBound  \triangleq\probability[\mathcal{D}]\prod_{x\in X}\sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D}]
\end{equation}
We alternate maximization of $\lowerBound$ w.r.t. $\left\{\mathcal{F}_x \right\}_{x\in X}$ %% (\textbf{Wake})
and $\mathcal{D}$:  %% (\textbf{Sleep-G})

\\\noindent \textbf{Wake: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $\mathcal{D}$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D}]$
is large.
Here we use a simple and generic enumerative program synthesis algorithm.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{Sleep-G: Maxing $ \lowerBound$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\mathcal{D}$.
To represent DSLs and search the space of DSLs, we use Fragment Grammars~\cite{tim}.

@@

\subsection{Programs that manipulate sequences}\label{sequences}
We apply \system to list processing %% (Section~\ref{listSection})
and text editing, %% (Section~\ref{textSection})
using a GRU~\cite{cho2014learning} for
the recognition model, and initially providing the system with generic sequence manipulation primitives:
\code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length},
\code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons},
\code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

%\subsubsection{List Processing}\label{listSection}
\textbf{List Processing:} Synthesizing programs that manipulate data structures is a widely studied
problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of learning functions that
manipulate lists, and also perform arithmetic operations upon lists (Table~\ref{listExamples}).
We created 236 human-interpretable list manipulation tasks, each with 15
input/output examples.
%% Our data set is interesting in three major ways: many of the tasks
%% require complex solutions; the tasks were not generated from some
%% latent DSL, and the agent must learn to solve these complicated
%% problems from only 236 tasks.
%% Our data set assumes arithmetic operations as well as sequence operations,
%% so we additionally provide our system with the following arithmetic
%% primitives: \code{mod}, \code{*}, \code{>}, \code{is-square},
%% \code{is-prime}.
%% evaluating  on a 50/50 test/train split.
In solving these tasks, the system
composed 38 new subroutines, and rediscovered the higher-order
function \code{filter} ($f_1$ in Table~\ref{initialExampleDSL}, left).%% , yielding a more expressive DSL more closely
%% matching the domain (Tbl.~\ref{initialExampleDSL}, left).
\begin{figure}[b]\centering
\vspace{-0.5cm}  \begin{tabular}{lll}
    \toprule
    Name & Input & Output \\\midrule
    repeat-3 & [7\, 0] & [7\, 0\, 7\, 0\, 7\, 0] \\
%    drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
    rotate-2 & [8\, 14\, 1\, 9] & [1\, 9\, 8\, 14] \\
%    count-head-in-tail & [1\, 2\, 1\, 1\, 3] & 2 \\
    keep-div-5 & [5\, 9\, 14\, 6\, 3\, 0] & [5\, 0] \\
    product & [7\, 1\, 6\, 2] & 84 \\
    \bottomrule
  \end{tabular}
  \captionof{table}{Some tasks in our list function domain.}\label{listExamples}\vspace{-0.5cm}
\end{figure}

@@

\subsection{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 %% symbolic regression
problems, each either a polynomial or rational function.  The recognition model is a
convnet that observes an image of the target function's
graph (Fig.~\ref{functions}) --- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the convnet can  look at a graph and predict
what kind of function best explains it.
These programs can contain real numbers,
which we set with gradient descent,
and penalize continuous parameters by
incorporating a BIC penalty into the likelihood model $\probability[x|p]$.
%% A key difficulty, however, is
%% that these problems are best solved with programs containing real
%% numbers.  Our solution to this difficulty is to enumerate
%%  programs with real-valued parameters, and then fit those
%% parameters by automatically differentiating through the programs the
%% system writes and use gradient descent to fit the parameters.
%% We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
%% and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
\begin{figure}\vspace{-0.0cm} \newcommand{\functionSize}{1.1cm}\centering
  \begin{minipage}[c]{0.3\columnwidth}%\vspace{-0.3cm}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}\\
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}%\includegraphics[width = \functionSize]{figures/functions/181.png}
  %\includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
    \includegraphics[width = \functionSize]{figures/functions/149.png}
  \end{minipage}
  \begin{minipage}[c]{0.69\columnwidth}    
    \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top rows) and rational functions (bottom) while the recognition  model jointly learns to look at a graph of the function (left) and predict which learned subroutines best explain the observation.}\label{functions}\vspace{-1cm}
        \end{minipage}
\end{figure}
We learn a DSL containing 13 new functions,
mainly templates for polynomials of different orders or ratios of polynomials.
The model also learns to find programs minimizing the number of continuous parameters ---
learning to represent linear functions with 
\code{(* real (+ x real))}, which has two continuous parameters, and represents quartic functions using $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing.

@@

We ran the following experiment: \system was given a subset of
the 1959 Lisp primitives, and tasked with solving 22
programming exercises.  After running for 93 hours on 64 CPUs, our
algorithm solves these exercises, along the way assembling a DSL
with a modern repertoire of
functional programming idioms and subroutines, including \code{map},
\code{fold}, \code{zip}, \code{unfold}, \code{index}, \code{length},
and  arithmetic operations like 
building lists of natural numbers between an interval.
%% We did not use the recognition model for this experiment:
%% a bottom-up pattern recognizer is of little use
%% for acquiring this abstract knowledge
%% from less than two dozen problems.

@@

\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models 
more closely matching the domain.

@@

\begin{figure*}\centering
  \includegraphics[width = 4.5cm]{figures/listLearningCurve.eps} \qquad
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}\qquad
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
\vspace{-0.4cm}  \caption{Learning curves for \system both with (\orange{in orange}) and without
    (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved w/ 10m timeout. Dashed lines: Average solve time, averaged only over tasks that are solved.}\label{learningCurves}\vspace{-0.5cm}
\end{figure*}

\section{Discussion}

%\paragraph{Outlook.}
We contribute an algorithm, \systemEnding, that learns to program by
bootstrapping a DSL with new domain-specific primitives that the algorithm
itself discovers, together with a neural recognition model that learns to
efficiently deploy the DSL on new tasks. %% We believe this integration of top-down
%% symbolic representations and bottom-up neural nets --- both of them learned
%% --- can help make program induction systems more generally useful for AI. 
%Many
%directions remain open.
%\paragraph{Future.}
Two immediate future goals are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the generic methods used here.
Another direction is DSL meta-learning: can we find a
\emph{single} universal primitive set that could bootstrap DSLs for
new domains, including the domains considered here,  but also many others?
